---
title: "AirlinePunctualityAnalysis"
author: "Stefan Huber"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)  
library(dplyr)  
library(ggplot2)  
```

### IMPORT
```{r}
airlines_data <- read.csv("airlines.csv") 
 
# Display the first few rows  
head(airlines_data)  
```

### TIDYING
```{r}
airlines_data$X <- ifelse(airlines_data$X == "", NA, airlines_data$X)
airlines_data <- airlines_data %>% fill(X)

# Rename columns for clarity
colnames(airlines_data) <- c("Airline", "Status", "Los_Angeles", "Phoenix", "San_Diego", "San_Francisco", "Seattle")
 
# Check the result  
head(airlines_data)  
```

### TRANSFORM
```{r}
airlines_data$Status <- ifelse(airlines_data$Status == "", NA, airlines_data$Status)

# Remove rows with NA in Status (these are likely header rows or empty rows)
airlines_data <- airlines_data %>% filter(!is.na(Status))

# Convert to long format
airlines_long <- airlines_data %>%
  pivot_longer(cols = c(Los_Angeles, Phoenix, San_Diego, San_Francisco, Seattle),
               names_to = "City",
               values_to = "Count")

# Check the long format data
print(head(airlines_long, 10))
```

### ANALYSIS
```{r}
overall_summary <- airlines_long %>%
  group_by(Airline, Status) %>%
  summarise(Total = sum(Count, na.rm = TRUE), .groups = 'drop')

overall_summary_wide <- overall_summary %>%
  pivot_wider(names_from = Status, values_from = Total) %>%
  mutate(
    Total_Flights = ifelse(is.na(`on time`), 0, `on time`) + ifelse(is.na(delayed), 0, delayed),
    pct_delayed = ifelse(Total_Flights > 0, 100 * ifelse(is.na(delayed), 0, delayed) / Total_Flights, 0)
  )

# Display overall summary
print(overall_summary_wide)

city_summary <- airlines_long %>%
  group_by(Airline, City, Status) %>%
  summarise(Total = sum(Count, na.rm = TRUE), .groups = 'drop')

city_summary_wide <- city_summary %>%
  pivot_wider(names_from = Status, values_from = Total) %>%
  mutate(
    Total_Flights = ifelse(is.na(`on time`), 0, `on time`) + ifelse(is.na(delayed), 0, delayed),
    pct_delayed = ifelse(Total_Flights > 0, 100 * ifelse(is.na(delayed), 0, delayed) / Total_Flights, 0)
  )

# Display city-level summary
print(city_summary_wide)

city_comparison <- city_summary_wide %>%
  select(Airline, City, pct_delayed) %>%
  rename(City_Pct_Delayed = pct_delayed)

overall_comparison <- overall_summary_wide %>%
  select(Airline, pct_delayed) %>%
  rename(Overall_Pct_Delayed = pct_delayed)

# Join the city and overall percentages
comparison_table <- city_comparison %>%
  left_join(overall_comparison, by = "Airline") %>%
  mutate(Difference = City_Pct_Delayed - Overall_Pct_Delayed)

# Display the comparison table
print(comparison_table)

# Calculate the average number of flights per city for each airline
flight_distribution <- airlines_long %>%
  group_by(Airline, City) %>%
  summarise(Total_Flights = sum(Count, na.rm = TRUE), .groups = 'drop')

# Display flight distribution
print(flight_distribution)
```

### VISUALIZE
```{r}
ggplot(city_summary_wide, aes(x = City, y = pct_delayed, fill = Airline)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "City-wise Percentage of Delayed Flights by Airline",
       x = "City",
       y = "Percentage Delayed (%)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### TEXT

### City-by-City:
### Los_Angeles: ALASKA performs better with 3.34% lower delay rate
### Phoenix: ALASKA performs better with 2.75% lower delay rate 
### San_Diego: ALASKA performs better with 5.89% lower delay rate
### San_Francisco: ALASKA performs better with 11.87% lower delay rate
### Seattle: ALASKA performs better with 9.07% lower delay rate

### ALASKA flight distribution: 
### Seattle: 56.85% of flights (2146 flights)
### San_Francisco: 16.03% of flights (605 flights)
### Los_Angeles: 14.81% of flights (559 flights)
### Phoenix: 6.17% of flights (233 flights)
### San_Diego: 6.15% of flights (232 flights)

### AM WEST flight distribution:
### Phoenix: 72.73% of flights (5255 flights)
### Los_Angeles: 11.22% of flights (811 flights)
### San_Francisco: 6.21% of flights (449 flights)
### San_Diego: 6.2% of flights (448 flights)
### Seattle: 3.63% of flights (262 flights)

### Simpson's Paradox:
### ALASKA performs better in every individual city, but as we can see from
### the data, AM WEST has better overall performance. 
### AM WEST has 72.73% of its flights in Phoenix (relatively low delay rate)
### ALASKA has most of its flights in Seattle (relatively higher delay rate)