---
title: "Bechel Test Analysis"
author: "Stefan Huber"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(fivethirtyeight)

bechdel_df <- bechdel
```

### OBJECTIVE

Exploring how to use TidyVerse packages to analyze the Bechdel test dataset (simple measure of female representation) from FiveThirtyEight.

### CLEAN

```{r data_prep}
bechdel_clean <- bechdel_df %>%
  select(year, title, binary, budget_2013, domgross_2013) %>%
  mutate(
    pass = ifelse(binary == "PASS", "Pass", "Fail"),
    roi = (domgross_2013 - budget_2013) / budget_2013
  )

head(bechdel_clean)
```

### ANALYZE

```{r pass_rates}
yearly_summary <- bechdel_clean %>%
  group_by(year) %>%
  summarize(
    total_movies = n(),
    pass_rate = mean(pass == "Pass"),
    avg_roi = mean(roi, na.rm = TRUE)
  )

ggplot(yearly_summary, aes(x = year, y = pass_rate)) +
  geom_line(color = "blue", linewidth = 1) +
  geom_smooth(method = "loess", se = TRUE, color = "red") +
  labs(
    title = "Movies Passing the Bechdel Test (1970-2013)",
    subtitle = "Percentage of movies that pass the Bechdel test has increased over time",
    x = "Year",
    y = "Pass Rate",
    caption = "Data source: FiveThirtyEight"
  ) +
  theme_minimal() +
  scale_y_continuous(labels = scales::percent)
```

### COMPARE

```{r budget_analysis}
bechdel_budget <- bechdel_df %>%
  filter(!is.na(budget_2013)) %>%
  mutate(pass = ifelse(binary == "PASS", "Pass", "Fail")) %>%
  group_by(pass) %>%
  summarize(
    avg_budget = mean(budget_2013, na.rm = TRUE),
    median_budget = median(budget_2013, na.rm = TRUE),
    count = n()
  )

print(bechdel_budget)
```

### ROI

```{r roi_analysis}
ggplot(yearly_summary, aes(x = year)) +
  geom_line(aes(y = pass_rate, color = "Pass Rate"), linewidth = 1) +
  geom_line(aes(y = avg_roi, color = "Average ROI"), linewidth = 1) +
  scale_color_manual(values = c("Pass Rate" = "blue", "Average ROI" = "red")) +
  labs(
    title = "Bechdel Test Pass Rate and ROI Over Time",
    x = "Year",
    y = "Rate",
    color = "Metric"
  ) +
  theme_minimal() +
  scale_y_continuous(labels = scales::percent)
```

### METHODS

- Cleaned and prepared data using dplyr
- Created visualizations with ggplot2
- Performed group-wise calculations
- Compared different metrics over time

### FINDINGS

- Pass rate for the Bechdel test has generally increased over time.
- Movies that fail the test tend to have higher budgets.