---
title: "Movie Ratings Analysis"
author: "Julius AI"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
This document demonstrates the process of creating a normalized SQLite database for movie ratings, loading the data into R, handling missing data, and performing analysis including standardizing ratings.

## Step 1: Create and Populate the Database
```{r create-database}
library(RSQLite)
library(tidyverse)

# Create a connection to a new SQLite database
con <- dbConnect(RSQLite::SQLite(), "movie_ratings.db")

# Create the Movies table
dbExecute(con, "
CREATE TABLE IF NOT EXISTS Movies (
    movie_id INTEGER PRIMARY KEY,
    title TEXT NOT NULL,
    release_year INTEGER
)")

# Create the Friends table
dbExecute(con, "
CREATE TABLE IF NOT EXISTS Friends (
    friend_id TEXT PRIMARY KEY,
    friend_name TEXT NOT NULL
)")

# Create the Ratings table
dbExecute(con, "
CREATE TABLE IF NOT EXISTS Ratings (
    rating_id INTEGER PRIMARY KEY AUTOINCREMENT,
    friend_id TEXT,
    movie_id INTEGER,
    rating INTEGER,
    FOREIGN KEY (friend_id) REFERENCES Friends(friend_id),
    FOREIGN KEY (movie_id) REFERENCES Movies(movie_id)
)")

# Insert movie data
movies_data <- data.frame(
    movie_id = 1:6,
    title = c("Dune", "Spider-Man: No Way Home", "The Batman", 
              "Top Gun: Maverick", "Avatar: The Way of Water", "Oppenheimer"),
    release_year = c(2021, 2021, 2022, 2022, 2022, 2023)
)

dbWriteTable(con, "Movies", movies_data, append = TRUE, row.names = FALSE)

# Insert friends data
friends_data <- data.frame(
    friend_id = c("F1", "F2", "F3", "F4", "F5"),
    friend_name = c("Friend A", "Friend B", "Friend C", "Friend D", "Friend E")
)

dbWriteTable(con, "Friends", friends_data, append = TRUE, row.names = FALSE)

# Insert ratings data
ratings_data <- data.frame(
    friend_id = c("F1", "F1", "F1", "F1", "F1", "F2", "F2", "F2", "F2", "F2", "F3", "F3", "F3", "F3", "F3", "F4", "F4", "F4", "F4", "F5", "F5", "F5", "F5", "F5"),
    movie_id = c(1, 2, 3, 4, 6, 1, 2, 3, 5, 6, 1, 3, 4, 5, 6, 2, 3, 4, 5, 1, 2, 4, 5, 6),
    rating = c(4, 5, 3, 4, 5, 5, 4, 4, 5, 4, 3, 5, 4, 4, 5, 5, 4, 5, 5, 4, 4, 3, 4, 5)
)

dbWriteTable(con, "Ratings", ratings_data, append = TRUE, row.names = FALSE)
```

## Step 2: Load Data into R
```{r load-data}
# Load the Ratings and Movies tables into R dataframes
ratings_df <- dbGetQuery(con, "SELECT * FROM Ratings")
movies_df <- dbGetQuery(con, "SELECT * FROM Movies")

# Display the first few rows of each dataframe
head(ratings_df)
head(movies_df)
```

## Step 3: Handle Missing Data
```{r handle-missing-data}
# Join ratings with movies and create a wide format table
ratings_wide <- dbGetQuery(con, "
    SELECT f.friend_id, m.title,
           CASE WHEN r.rating IS NULL THEN 'NA' ELSE CAST(r.rating AS TEXT) END as rating
    FROM Friends f
    CROSS JOIN Movies m
    LEFT JOIN Ratings r ON f.friend_id = r.friend_id AND m.movie_id = r.movie_id
    ORDER BY f.friend_id, m.movie_id
")

# Convert to wide format using tidyr
ratings_matrix <- ratings_wide %>%
    pivot_wider(names_from = title, values_from = rating) %>%
    as.data.frame()

# Display the ratings matrix
ratings_matrix
```

## Step 4: Standardize Ratings
```{r standardize-ratings}
# Convert ratings to numeric and calculate z-scores for each movie
ratings_standardized <- dbGetQuery(con, "
    WITH MovieStats AS (
        SELECT movie_id,
               AVG(CAST(rating AS FLOAT)) as mean_rating,
               SQRT(AVG(CAST(rating AS FLOAT) * CAST(rating AS FLOAT)) - 
                    AVG(CAST(rating AS FLOAT)) * AVG(CAST(rating AS FLOAT))) as std_dev
        FROM Ratings
        GROUP BY movie_id
    )
    SELECT r.friend_id,
           m.title,
           r.rating as original_rating,
           ROUND(CAST((r.rating - ms.mean_rating) / CASE WHEN ms.std_dev = 0 THEN 1 ELSE ms.std_dev END AS FLOAT), 2) as z_score
    FROM Ratings r
    JOIN Movies m ON r.movie_id = m.movie_id
    JOIN MovieStats ms ON r.movie_id = ms.movie_id
    ORDER BY r.friend_id, m.movie_id
")

ratings_standardized
```

## Step 5: Summary Statistics
```{r summary-statistics}
# Calculate summary statistics
rating_summary <- dbGetQuery(con, "
    SELECT m.title,
           COUNT(r.rating) as num_ratings,
           ROUND(AVG(CAST(r.rating AS FLOAT)), 2) as avg_rating,
           MIN(r.rating) as min_rating,
           MAX(r.rating) as max_rating
    FROM Movies m
    LEFT JOIN Ratings r ON m.movie_id = r.movie_id
    GROUP BY m.title
    ORDER BY m.movie_id
")

rating_summary
```

## Conclusion
This analysis demonstrates the process of creating a normalized database, handling missing data, and performing statistical analysis in R. The SQLite database and R code ensure reproducibility and scalability for similar projects.

```{r close-connection, include=FALSE}
# Close the database connection
dbDisconnect(con)
```

