---
title: 'RegEx Analysis'
author: 'Stefan Huber'
output: html_document
---

### 1. Using the 173 majors listed in fivethirtyeight.com’s College Majors 
### dataset, provide code that identifies the majors that contain either 
### "DATA" or "STATISTICS

```{r}
library(dplyr)

majors <- read.csv('https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv')

data_statistics_majors <- majors %>% 
  filter(grepl('DATA', Major, ignore.case = TRUE) | 
         grepl('STATISTICS', Major, ignore.case = TRUE))

print(data_statistics_majors)
```

### 2 Write code that transforms the data below:

### [1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

### [5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

### [9] "elderberry"   "lime"         "lychee"       "mulberry"    

### [13] "olive"        "salal berry"

### Into a format like this:

### c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", 
### "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee",
### "mulberry", "olive", "salal berry")

```{r}
fruits <- c('bell pepper', 'bilberry', 'blackberry', 'blood orange',
            'blueberry', 'cantaloupe', 'chili pepper', 'cloudberry',
            'elderberry', 'lime', 'lychee', 'mulberry', 'olive', 'salal berry')

print(fruits)
```

### The two exercises below are taken from R for Data Science, 
### 14.3.5.1 in the on-line version:

### 3 Describe, in words, what these expressions will match:

### (.)\1\1
### "(.)(.)\\2\\1"
### (..)\1
### "(.).\\1.\\1"
### "(.)(.)(.).*\\3\\2\\1"

### (.)\1\1
### Three of the same character in sequence
### I.e. 'aaa'

### "(.)(.)\\2\\1"
### Two letters that repeat in reverse
### I.e. 'abba' (like the band)

### (..)\1
### Two letters that repeat in sequence
### I.e. 'abab'

### "(.).\\1.\\1"
### A letter -> any letter -> the first letter -> any letter > first letter
### I.e. 'aXaYa'

### "(.)(.)(.).*\\3\\2\\1"
### Three letters -> anything -> those three letters backwards
### I.e. 'abcXYZcba'

### 4 Construct regular expressions to match words that:

### Start and end with the same character.
### Contain a repeated pair of letters
### (e.g. "church" contains "ch" repeated twice.)
### Contain one letter repeated in at least three places
### (e.g. "eleven" contains three "e"s.)

### Words that start and end with the same letter:
```{r}
# The ^ means start, $ means end
regex1 <- '^(.).*\1$'
print(regex1)  # Try it with words like 'anna' or 'eye'
```

### Words with a repeated pair of letters:
```{r}
# Looking for any two letters that show up twice
regex2 <- '(..).*\1'
print(regex2)  # Works with words like 'murmur'
```

### Words with one letter showing up at least three times:
```{r}
# Finding a letter that appears three times anywhere
regex3 <- '.*([A-Za-z]).*\1.*\1.*'
print(regex3)  # Try it with 'banana'
```
