---
title: "Untitled"
author: "Stefan Huber"
output: html_document
---


### OBJECTIVE

##  Reading and cleaning three CSV files from last week's Discussion:  
## (1) NYC Gifted and Talented Grades 2018-19 (Sheet5)
## (2) D&D Characters All
## (3) Untidy Dataset _624

```{r setup, include=FALSE}
library(tidyverse)
```

### (1) READING

```{r}
nyc_gt <- read.csv("NYC Gifted and Talented Grades 2018-19 - Sheet5.csv", stringsAsFactors = FALSE)  
 
head(nyc_gt)  
str(nyc_gt)
```
  
### (1) TIDYING

```{r}
nyc_gt <- nyc_gt %>%  
 rename(  
   Timestamp = Timestamp,  
   GradeLevel = Entering.Grade.Level,  
   District = District,  
   BirthMonth = Birth.Month,  
   OLSAT_Verbal_Score = OLSAT.Verbal.Score,  
   OLSAT_Verbal_Percentile = OLSAT.Verbal.Percentile,  
   NNAT_NonVerbal_Raw = NNAT.Non.Verbal.Raw.Score,  
   NNAT_NonVerbal_Percentile = NNAT.Non.Verbal.Percentile,  
   Overall_Score = Overall.Score,  
   School_Preferences = School.Preferences,  
   Extra = X,  
   School_Assigned = School.Assigned,  
   Enroll = Will.you.enroll.there.  
 )  

nyc_gt_tidy <- nyc_gt %>% filter(School_Assigned != "")  
 
head(nyc_gt_tidy)  
```

### (1) ANALYSIS

```{r}
summary(nyc_gt_tidy$OLSAT_Verbal_Score)  

enroll_summary <- nyc_gt_tidy %>%  
 group_by(Enroll) %>%  
 summarise(count = n())  
print(enroll_summary)  
 
library(ggplot2)  
ggplot(nyc_gt_tidy, aes(x = OLSAT_Verbal_Score)) +  
 geom_histogram(binwidth = 1, fill = "blue", color = "black") +  
 labs(title = "Distribution of OLSAT Verbal Scores", x = "Score", y = "Count")  
```

### (2) READING

```{r}
dnd_chars <- read.csv("dnd_chars_all.csv", stringsAsFactors = FALSE)  
 
head(dnd_chars)  
str(dnd_chars)  
```

### (2) TIDYING

```{r}
if("date" %in% names(dnd_chars)){  
 dnd_chars$date <- as.Date(dnd_chars$date)  
}  
 
dnd_chars_tidy <- dnd_chars %>%  
 mutate(  
   Primary_Class = sapply(strsplit(as.character(class), '\\\\|'), `[`, 1),  
   Secondary_Class = sapply(strsplit(as.character(class), '\\\\|'), function(x) ifelse(length(x)>1, x[2], NA))  
 )  
 
head(dnd_chars_tidy) 
```

### (2) ANALYSIS

```{r}
summary(dnd_chars_tidy$level)  
 
class_counts <- dnd_chars_tidy %>% group_by(Primary_Class) %>% summarise(count = n())  
print(class_counts)  
 
ggplot(dnd_chars_tidy, aes(x = level)) +  
 geom_histogram(binwidth = 1, fill = "green", color = "black") +  
 labs(title = "Distribution of Character Levels", x = "Level", y = "Count")  
```

### (3) READING

```{r}
untidy <- read.csv("Untidydataset _624.csv", stringsAsFactors = FALSE)  
 
head(untidy)  
str(untidy) 
```

### (3) TIDYING

```{r}
untidy <- untidy %>%   
 rename_all(~ make.names(.))  
 
 
if(all(c('measurement1','measurement2','measurement3') %in% names(untidy))){  
 untidy_tidy <- untidy %>%   
   pivot_longer(cols = c(measurement1, measurement2, measurement3),   
                names_to = "Measurement", values_to = "Value")  
} else {  
 untidy_tidy <- untidy # if no such columns exist, use the data as is.  
}  
 
head(untidy_tidy)  
```

### (3) ANALYSIS

```{r}
summary(untidy_tidy)  
 
if("Value" %in% names(untidy_tidy)){  
 ggplot(untidy_tidy, aes(x = Value)) +  
   geom_histogram(fill = "purple", color = "black") +  
   labs(title = "Distribution of Values", x = "Value", y = "Count")  
}  
```

### CONCLUSION

## (1) NYC Gifted and Talented Data
## Removed rows with missing school assignments
## Analyzed test score distributions

## (2) D&D Characters Data
## Converted dates
## Split combined class data
## Analyzed character levels and class distributions

## (3) Untidy Dataset _624
## Renamed columns and potentially gathered measurement columnns
## Summarized dataset with descriptive statistics and a plot